### 2023-01-05 Code/Simulation.Rmd
Autoarma histogram framework with ggplot

---
title: "R Notebook"
output: html_notebook
---

```{r packages}
# Load packages
library(tidyverse)
library(copula)
library(evd)
library(ismev)
library(tseries)
library(forecast)
source('tspb.R')
```

```{r hist_correlation}
# Test to show that autocorrelation changes the power of the test
test_correlation <- function(n_tests, ars) {

  # Create a master data_frame for p-values
  columns <- c("p_value", "ar")
  ar_results <- data.frame(matrix(ncol = 2, nrow = 0))
  colnames(ar_results) <- columns

  # For AR coefficients from -0.5 to 0.5
  for(i in 1:length(ars)) {

    for(j in 1:n_tests) {

      # Generate random correlated data, perform a KS-test, and store the p-value
      data <- arima.sim(model = list(ar = ars[i]), rand.gen = rnorm, sd = sqrt(1-ars[i]^2), n = 100)
      ar_results <- add_row(ar_results, p_value = ks.test(data, "pnorm")$p.value, ar = round(ars[i], 1))
    }
  }

  return(ar_results)
}

 # Set AR coefficients and generate results
ars <- seq(-0.3, 0.3, by = 0.1)
ar_results <- test_correlation(1000, ars)

# Generate histograms plot
hist_correlation <- ggplot(ar_results) + 
                    geom_histogram(aes(x = p_value, fill = ar), boundary = 0, binwidth = 1/20) + 
                    facet_grid(~ar) +
                    theme(legend.position = "none") +
                    scale_x_continuous(name = "p-value", breaks=c(0, 1))

# Save the histogram figure
ggsave(filename = "hist_correlation.pdf", plot = hist_correlation, path = "../Manuscript", height = 2.1, width = 6)
```

```{r Parametric}
# Test parametric bootstrap as a solution to fitting parameters
test_parametric <- function(x, B = 1000) {

  # Summarize parameters
  mu <- mean(x)
  sigma <- sd(x)

  # Store KS statistic from using fitted parameters
  stat <- ks.test(x, "pnorm", mu, sigma)$statistic

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rnorm(n, mu, sigma)
    mu.b <- mean(x.b)
    sigma.b <- sd(x.b)
    stat.b[i] <- ks.test(x.b, "pnorm", mu.b, sigma.b)$statistic
  }

  # Return empirical p-value
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

# Test non-parametric bootstrap as a solution to fitted parameters
test_babu <- function(x, B = 1000) {
  p.value <- ks.np.babu(x, B = 1000)$p.value
  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

# Naive KS-test to show consequence of fitted parameters
test_naive <- function(x) {
  p.value <- ks.test(x, "pnorm", mean = mean(x), sd = sd(x))$p.value
  list(p.value = p.value)
}

# Generate results
n_tests <- 1000
n <- 100
counter <- 0
pvals_parametric <- replicate(n_tests, test_parametric(rnorm(n), B = 1000)$p.value)
counter <- 0
pvals_babu <- replicate(n_tests, test_babu(rnorm(n), B = 1000)$p.value)
pvals_naive <- replicate(n_tests, test_naive(rnorm(n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Parametric", n_tests), rep("Babu", n_tests), rep("Naive", n_tests))
p_values <- c(pvals_parametric, pvals_babu, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Parametric", "Babu"))

# Generate histograms plot
hist_fitted <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

# Save the histogram figure
ggsave(filename = "hist_fitted.pdf", plot = hist_fitted, path = "../Manuscript", height = 2.1, width = 6)
```

```{r babu}
ecdf.leftcont <- function (x) {
    x <- sort(x)
    n <- length(x)
    if (n < 1) 
        stop("'x' must have 1 or more non-missing values")
    vals <- unique(x)
    ## making it left continuous!
    rval <- approxfun(vals, cumsum(tabulate(match(x, vals)))/n - 1/n, 
        method = "constant", yleft = 0, yright = 1, f = 1, ties = "ordered")
    class(rval) <- c("ecdf", "stepfun", class(rval))
    assign("nobs", n, envir = environment(rval))
    attr(rval, "call") <- sys.call()
    rval
}

ks.np.babu <- function(x, B = 1000) {
    ## for normal distribution only
    n <- length(x)
    Fn <- ecdf(x)
    Gn <- ecdf.leftcont(x)
    mu <- mean(x); ss <- sd(x)
    Ftheta <- pnorm(x, mu, ss)
    Bn1 <- Fn(x) - Ftheta
    Bn2 <- Gn(x) - Ftheta
    d1 <- abs(Bn1)
    d2 <- abs(Bn2)
    stat <- max(d1, d2)
    stat.b <- double(B)
    for (i in 1:B) {
        x.b <- sample(x, size = n, replace = TRUE)
        Fn.b <- ecdf(x.b)
        Gn.b <- ecdf.leftcont(x.b)
        mu.b <- mean(x.b); ss.b <- sd(x.b)
        ## evaluate at x (not x.b)
        Ftheta.b <- pnorm(x, mu.b, ss.b)
        d1.b <- abs(Fn.b(x) - Ftheta.b - Bn1)
        d2.b <- abs(Gn.b(x) - Ftheta.b - Bn2)
        stat.b[i] <- max(d1.b, d2.b)
    }
    p.value <-  (sum(stat.b >= stat) + 0.5) / (B + 1)
    list(p.value = p.value,
         statistic = stat, stat.b = stat.b)
}
```

```{r gen_data}
## generating observations from a specified distribution with serial dependence
genData <- function(n, qdist, ..., copula, rho) {
    ## n: sample size
    ## qdist: quantile function of the desired marginal distribution
    ## copula: copula for serial distribution
    ## rho: spearman's rho
    theta <- iRho(copula(), rho)
    cop <- copula(theta)
    u <- runif(n)
    for (j in 2:n) {
        u[j] <- cCopula(c(u[j-1], u[j]), cop, indices = 2, inverse = TRUE)
    }
    qdist(u, ...)
}
```

```{r AR(1) D}
test_copula_D <- function(x, pdist, qdist, ..., B, wcopula = normalCopula) {

    # Get observed statistic
    n <- length(x)

    stat  <- ks.test(x, pdist, ...)$statistic

    # Get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(wcopula(), rho)
    wcop <- wcopula(r)

    # Perform parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)

    for (i in 1:B) {
        # 1. The marginal distribution is the fitted dist
        # 2. There is temporal dependence
        u <- runif(n) 
        for (j in 2:n) {
            u[j]  <- cCopula(c(u[j-1], u[j]), wcop, indices = 2, inverse = TRUE)
        }

        # Transform bootstrap sample and fit the parameters
        x.b <- qdist(u, ...)

        # Gather bootstrap sample's test statistic
        stat.b[i] <- ks.test(x.b, pdist, ...)$statistic
    }

    # Return empirical p-value
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

    counter <<- counter + 1
    cat(counter, "")

    return(p.value)
}

# Standard KS test
test_standard_ks <- function(x, pdist, ...) {
  p.value <- ks.test(x, pdist, ...)$p.value
  list(p.value = p.value)
}

ptm <- proc.time()

# Set variables
n <- 100
ar <- 0.5
counter <- 0
nrep <- 1000
B <- 100
mean <- 0
sd <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(genData(n = 100, qdist = qnorm, mean = 0, sd = 1, copula = claytonCopula, rho = 0.5), 
                                              pdist = pnorm, 
                                              qdist = qnorm,
                                              mean = mean,
                                              sd = sd,
                                              B = B, 
                                              wcopula = normalCopula))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(genData(n = 100, qdist = qnorm, mean = 0, sd = 1, copula = claytonCopula, rho = 0.5), 
                                                pdist = pnorm,
                                                mean = mean, 
                                                sd = sd)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_ar1_D <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_ar1_D

print(proc.time() - ptm)

# Save the histogram figure
ggsave(filename = "hist_ar1_D.pdf", plot = hist_ar1_D, path = "../Manuscript", height = 2.1, width = 6)
```

```{r MA(1) D}
ptm <- proc.time()

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 1000
ma <- 0.5
mean <- 0
sd <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(arima.sim(list(ma = 0.5), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n), 
                                              pdist = pnorm, 
                                              qdist = qnorm,
                                              mean = mean,
                                              sd = sd,
                                              B = B,
                                              wcopula = normalCopula))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(arima.sim(list(ma = 0.5), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n),
                                                pdist = pnorm,
                                                mean = mean,
                                                sd = sd)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_ma1_D <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_ma1_D

print(proc.time() - ptm)

# Save the histogram figure
ggsave(filename = "hist_ma1_D.pdf", plot = hist_ma1_D, path = "../Manuscript", height = 2.1, width = 6)
```

```{r ARMA(1,1) D}
ptm <- proc.time()

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 1000
ar <- 0.5
ma <- 0.3
mean <- 0
sd <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(arima.sim(list(ar = ar, 
                                                           ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n), 
                                            pdist = pnorm, 
                                            qdist = qnorm,
                                            mean = mean,
                                            sd = sd,
                                            B = B,
                                            wcopula = normalCopula))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(arima.sim(list(ar = ar, 
                                                           ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n),
                                                pdist = pnorm,
                                                mean = mean,
                                                sd = sd)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_arma_D <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_arma_D

print(proc.time() - ptm)

# Save the histogram figure
ggsave(filename = "hist_arma_D.pdf", plot = hist_arma_D, path = "../Manuscript", height = 2.1, width = 6)
```

```{r AR(2) D}
ptm <- proc.time()

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 1000
b1 <- 0.5
b2 <- 0.3
mean <- 0
sd <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(arima.sim(list(order = c(2, 0, 0), 
                                                           ar = c(b1, b2)),
                                                      rand.gen=rnorm,
                                                      sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)),
                                                      n = n), 
                                            pdist = pnorm, 
                                            qdist = qnorm,
                                            mean = mean,
                                            sd = sd,
                                            B = B,
                                            wcopula = normalCopula))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(arima.sim(list(order = c(2, 0, 0), 
                                                           ar = c(b1, b2)),
                                                      rand.gen=rnorm,
                                                      sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)),
                                                      n = n),
                                                pdist = pnorm,
                                                mean = mean,
                                                sd = sd)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_ar2_D <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_ar2_D

print(proc.time() - ptm)

# Save the histogram figure
ggsave(filename = "hist_ar2_D.pdf", plot = hist_ar2_D, path = "../Manuscript", height = 2.1, width = 6)
```

```{r Gamma D}
ptm <- proc.time()

# Set variables
n <- 100
rho <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
shape <- 3
scale <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(genData(n, qgamma, shape = shape, scale = scale, copula = normalCopula, rho = rho), 
                                             pdist = pgamma,
                                             qdist = qgamma,
                                             shape = shape,
                                             scale = scale,
                                             B = B,
                                             wcopula = normalCopula))
counter <- 0

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_standard_ks(genData(n, qgamma, shape = shape, scale = scale, copula = normalCopula, rho = rho), 
                                                   pdist = pgamma,
                                                   shape = shape, 
                                                   scale = scale)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_gamma_D <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_gamma_D

print(proc.time() - ptm)

# Save the histogram figure
ggsave(filename = "hist_gamma_D.pdf", plot = hist_gamma_D, path = "../Manuscript", height = 2.1, width = 6)
```

```{r GEV D}
ptm <- proc.time()

# Set variables
n <- 100
rho <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
loc <- 0
shape <- 0.2
scale <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(genData(n, qgev, loc = loc, shape = shape, scale = scale, copula = normalCopula, rho = rho),
                                           pdist = pgev,
                                           qdist = qgev,
                                           loc = loc, 
                                           shape = shape,
                                           scale = scale,
                                           B = B,
                                           wcopula = normalCopula))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(genData(n, qgev, loc = loc, shape = shape, scale = scale, copula = normalCopula, rho = rho), 
                                                 pdist = pgev,
                                                 loc = loc, 
                                                 shape = shape,
                                                 scale = scale)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_gev_D <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_gev_D

print(proc.time() - ptm)

# Save the histogram figure
ggsave(filename = "hist_gev_D.pdf", plot = hist_gev_D, path = "../Manuscript", height = 2.1, width = 6)
```

```{r AR(1) FD}
test_copula_FD <- function(x, dist, B) {

    # Get observed statistic
    n <- length(x)
    mu <- mean(x)
    sigma <- sd(x)
    stat  <- ks.test(x, dist, mean = mu, sd = sigma)$statistic

    # Get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(normalCopula(), rho)
    np <- normalCopula(r)

    # Perform parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)

    for (i in 1:B) {
        # 1. The marginal distribution is the fitted dist
        # 2. There is temporal dependence
        u <- runif(n) 
        for (j in 2:n) {
            u[j]  <- cCopula(c(u[j-1], u[j]), np, indices = 2, inverse = TRUE)
        }

        # Transform bootstrap sample and fit the parameters
        x.b <- qnorm(u, mean = mu, sd = sigma)
        mu.b <- mean(x.b)
        sigma.b <- sd(x.b)

        # Gather bootstrap sample's test statistic
        stat.b[i] <- ks.test(x.b, dist, mean = mu.b, sd = sigma.b)$statistic
    }

    # Return empirical p-value
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

    counter <<- counter + 1
    cat(counter, "")

    return(p.value)
}

# Super naive KS-test with fitted parameters
test_standard_ks_F <- function(x) {
  p.value <- ks.test(x, "pnorm", mean = mean(x), sd = sd(x))$p.value
  list(p.value = p.value)
}

ptm <- proc.time()

# Set variables
n <- 100
ar <- 0.5
counter <- 0
nrep <- 1000
B <- 1000

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_FD(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n), "pnorm", B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_standard_ks_F(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Super Naive", "Naive Parametric", "Copula"))

# Generate histograms plot
hist_ar1_FD <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_ar1_FD

print(proc.time() - ptm)

# Save the histogram figure
ggsave(filename = "hist_ar1_FD.pdf", plot = hist_ar1_FD, path = "../Manuscript", height = 2.1, width = 6)
```

```{r MA(1) FD}
## Test MA(1) model
ptm <- proc.time()

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 1000
ma <- 0.5

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(arima.sim(list(ma = ma), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n), "pnorm", B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric(arima.sim(list(ma = ma), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_standard_ks_F(arima.sim(list(ma = ma), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Super Naive", "Naive Parametric", "Copula"))

# Generate histograms plot
hist_ma1_FD <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

print(proc.time() - ptm)

hist_ma1_FD

# Save the histogram figure
ggsave(filename = "hist_ma1_FD.pdf", plot = hist_ma1_FD, path = "../Manuscript", height = 2.1, width = 6)
```

```{r ARMA(1,1) FD}
## Test ARMA(1,1) model
ptm <- proc.time()

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 1000
ar <- 0.5
ma <- 0.3

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(arima.sim(list(ar = ar, 
                                                           ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n),
                                            "pnorm", B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric(arima.sim(list(ar = ar, 
                                                              ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_standard_ks_F(arima.sim(list(ar = ar, 
                                                              ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Super Naive", "Naive Parametric", "Copula"))

# Generate histograms plot
hist_arma_FD <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

print(proc.time() - ptm)

hist_arma_FD

# Save the histogram figure
ggsave(filename = "hist_arma_FD.pdf", plot = hist_arma_FD, path = "../Manuscript", height = 2.1, width = 6)
```

```{r AR(2) FD}
## Test AR(2) model
ptm <- proc.time()

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 1000
b1 <- 0.5
b2 <- 0.3

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D(arima.sim(list(order = c(2, 0, 0), 
                                                           ar = c(b1, b2)),
                                                      rand.gen=rnorm,
                                                      sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)),
                                                      n = n), 
                                            "pnorm", 
                                            B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric(arima.sim(list(order = c(2, 0, 0), 
                                                                   ar = c(b1, b2)), 
                                                              rand.gen=rnorm, 
                                                              sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)), 
                                                              n = n))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_super_naive(arima.sim(list(order = c(2, 0, 0), 
                                                               ar = c(b1, b2)), 
                                                          rand.gen=rnorm, 
                                                          sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)), 
                                                          n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Super Naive", "Naive Parametric", "Copula"))

# Generate histograms plot
hist_ar2_FD <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

print(proc.time() - ptm)

hist_ar2_FD

# Save the histogram figure
ggsave(filename = "hist_ar2_FD.pdf", plot = hist_ar2_FD, path = "../Manuscript", height = 2.1, width = 6)
```

```{r Gamma FD}
# Testing bootstrap with copulas for gamma distribution
test_gamma_FD <- function(x, dist, B = 100, wcopula = normalCopula) {

    # Get observed stat
    n <- length(x)
    m <- mean(x)
    v <- var(x)
    scale <- v/m
    shape <- m*m/v

    stat  <- ks.test(x, dist, shape = shape, scale = scale)$statistic

    # Get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(wcopula(), rho)
    wcop <- wcopula(r)

    # Parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)

    for (i in 1:B) {
        # 1. The marginal distribution is the fitted dist
        # 2. There is temporal dependence
        u <- runif(n)
        for (j in 2:n) {
            u[j]  <- cCopula(c(u[j-1], u[j]), wcop, indices = 2, inverse = TRUE)
        }

        # Transform the bootstrap sample and fit the parameters
        x.b <- qgamma(u, shape = shape, scale = scale)

        m.b <- mean(x.b)
        v.b <- var(x.b)
        scale.b <- v.b/m.b
        shape.b <- m.b*m.b/v.b

        # Calculate new test statistic
        stat.b[i] <- ks.test(x.b, dist, shape = shape.b, scale = scale.b)$statistic
    }

    # Return empirical p-value    
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

    counter <<- counter + 1
    cat(counter, "")

    return(p.value)
}

# Parametric bootstrap to solve fitted parameters for gamma distribution
test_parametric_gamma <- function(x, B = 1000) {

  # Summarize parameters
  m <- mean(x)
  v <- var(x)
  scale <- v/m
  shape <- m*m/v

  # Store KS statistic from using fitted parameters
  stat <- ks.test(x, "pgamma", shape = shape, scale = scale)$statistic

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rgamma(n, shape = shape, scale = scale)

    m.b <- mean(x.b)
    v.b <- var(x.b)

    scale.b <- v.b/m.b
    shape.b <- m.b*m.b/v.b

    stat.b[i] <- ks.test(x.b, "pgamma", shape = shape.b, scale = scale.b)$statistic
  }

  # Return empirical p-value
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

# Naive function that incorrectly uses fitted parameters for gamma distribution
test_super_naive_gamma <- function(x) {

  m <- mean(x)
  v <- var(x)
  scale <- v/m
  shape <- m*m/v

  p.value <- ks.test(x, "pgamma", shape = shape, scale = scale)$p.value
  list(p.value = p.value)
}

ptm <- proc.time()

# Set variables
n <- 100
rho <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
shape <- 3
scale <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_gamma_FD(genData(n, qgamma, shape = shape, scale = scale, copula = normalCopula, rho = rho), "pgamma", B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric_gamma(genData(n, qgamma, shape = shape, scale = scale, copula = normalCopula, rho = rho))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_super_naive_gamma(genData(n, qgamma, shape = shape, scale = scale, copula = normalCopula, rho = rho))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Super Naive", "Naive Parametric", "Copula"))

# Generate histograms plot
hist_gamma_FD <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

print(proc.time() - ptm)

hist_gamma_FD

# Save the histogram figure
ggsave(filename = "hist_gamma_FD.pdf", plot = hist_gamma_FD, path = "../Manuscript", height = 2.1, width = 6)
```

```{r GEV FD}
# Testing bootstrap with copulas for GEV distribution
test_gev_FD <- function(x, B = 100, wcopula = normalCopula) {

    n <- length(x)

    # Get observed stat
    fit <- gev.fit(x, show = FALSE)
    loc <- fit$mle[1]
    scale <- fit$mle[2]
    shape <- fit$mle[3]

    stat  <- ks.test(x, 'pgev', loc = loc, scale = scale, shape = shape)$statistic

    # Get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(wcopula(), rho)
    wcop <- wcopula(r)

    # Parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)

    for (i in 1:B) {
        # 1. The marginal distribution is the fitted dist
        # 2. There is temporal dependence
        u <- runif(n)
        for (j in 2:n) {
            u[j]  <- cCopula(c(u[j-1], u[j]), wcop, indices = 2, inverse = TRUE)
        }

        # Transform the bootstrap sample and fit the parameters
        x.b <- qgev(u, loc = loc, scale = scale, shape = shape)

        fit.b <- gev.fit(x.b, show = FALSE)
        loc.b <- fit.b$mle[1]
        scale.b <- fit.b$mle[2]
        shape.b <- fit.b$mle[3]

        # Calculate bootstrap test statistic
        stat.b[i] <- ks.test(x.b, 'pgev', loc = loc.b, scale = scale.b, shape = shape.b)$statistic
    }

    # Return empirical p-value    
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

    counter <<- counter + 1
    cat(counter, "")

    return(p.value)
}

# Parametric bootstrap to solve fitted parameters for GEV distribution
test_parametric_gev <- function(x, B = 1000) {

  # Summarize parameters
  fit <- gev.fit(x, show = FALSE)
  loc <- fit$mle[1]
  scale <- fit$mle[2]
  shape <- fit$mle[3]

  # Store KS statistic from using fitted parameters
  stat <- ks.test(x, "pgev", loc = loc, scale = scale, shape = shape)$statistic

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rgev(n, loc = loc, scale = scale, shape = shape)

    fit.b <- gev.fit(x.b, show = FALSE)
    loc.b <- fit.b$mle[1]
    scale.b <- fit.b$mle[2]
    shape.b <- fit.b$mle[3]

    stat.b[i] <- ks.test(x.b, "pgev", loc.b, shape = shape.b, scale = scale.b)$statistic
  }

  # Return empirical p-value
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

# Naive function that incorrectly uses fitted parameters for GEV distribution
test_super_naive_gev <- function(x) {

  fit <- gev.fit(x, show = FALSE)
  loc <- fit$mle[1]
  scale <- fit$mle[2]
  shape <- fit$mle[3]

  p.value <- ks.test(x, "pgev", loc = loc, scale = scale, shape = shape)$p.value
  list(p.value = p.value)
}

ptm <- proc.time()

# Set variables
n <- 100
rho <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
loc <- 0
shape <- 0.2
scale <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_gev_FD(genData(n, qgev, loc = loc, shape = shape, scale = scale, copula = normalCopula, rho = rho), B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- suppressWarnings(replicate(nrep, test_parametric_gamma(genData(n, qgev, loc = loc, shape = shape, scale = scale, copula = normalCopula, rho = rho))$p.value))

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_super_naive_gamma(genData(n, qgev, loc = loc, shape = shape, scale = scale, copula = normalCopula, rho = rho))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Super Naive", "Naive Parametric", "Copula"))

# Generate histograms plot
hist_gev_FD <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

print(proc.time() - ptm)

hist_gev_FD

# Save the histogram figure
ggsave(filename = "hist_gev_FD.pdf", plot = hist_gev_FD, path = "../Manuscript", height = 2.1, width = 6)
```

```{r hist_ma1_arma_ar2_D}
load('results_ma1_D.RData')
model <- rep("MA(1)", 2000)
ma1 <- data_frame(results, model)

load('results_arma_D.RData')
model <- rep("ARMA(1, 1)", 2000)
arma <- data_frame(results, model)

load('results_ar2_D.RData')
model <- rep("AR(2)", 2000)
ar2 <- data_frame(results, model)

data <- rbind(ma1, arma, ar2)

data$model <- factor(data$model, levels = c("MA(1)", "ARMA(1, 1)", "AR(2)"))
data$histogram <- factor(data$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_ma1_arma_ar2_D <- ggplot(data) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(histogram~model, scales = "free_y")
hist_ma1_arma_ar2_D

ggsave(filename = "hist_ma1_arma_ar2_D.pdf", plot = hist_ma1_arma_ar2_D, path = "../Manuscript", height = 4.2, width = 6)
```

```{r hist_gamma_gev_D}
load('results_gamma_D.RData')
model <- rep("Gamma", 2000)
gamma <- data_frame(results, model)

load('results_gev_D.RData')
model <- rep("GEV", 2000)
gev <- data_frame(results, model)

data <- rbind(gamma, gev)

data$model <- factor(data$model, levels = c("Gamma", "GEV"))
data$histogram <- factor(data$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_gamma_gev_D <- ggplot(data) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(histogram~model, scales = "free_y")
hist_gamma_gev_D

ggsave(filename = "hist_gamma_gev_D.pdf", plot = hist_gamma_gev_D, path = "../Manuscript", height = 4.2, width = 6)
```

```{r hist_ma1_arma_ar2_FD}
load('results_ma1_FD.RData')
model <- rep("MA(1)", 3000)
ma1 <- data_frame(results, model)

load('results_arma_FD.RData')
model <- rep("ARMA(1, 1)", 3000)
arma <- data_frame(results, model)

load('results_ar2_FD.RData')
model <- rep("AR(2)", 3000)
ar2 <- data_frame(results, model)

data <- rbind(ma1, arma, ar2)

data$model <- factor(data$model, levels = c("MA(1)", "ARMA(1, 1)", "AR(2)"))
data$histogram <- factor(data$histogram, levels = c("Super Naive", "Naive Parametric", "Copula"))

# Generate histograms plot
hist_ma1_arma_ar2_FD <- ggplot(data) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_grid(histogram~model, scales = "free_y")
hist_ma1_arma_ar2_FD

ggsave(filename = "hist_ma1_arma_ar2_FD.pdf", plot = hist_ma1_arma_ar2_FD, path = "../Manuscript", height = 4.2, width = 6)
```

```{r hist_gamma_gev_FD}
load('results_gamma_FD.RData')
model <- rep("Gamma", 3000)
gamma <- data_frame(results, model)

load('results_gev_FD.RData')
model <- rep("GEV", 3000)
gev <- data_frame(results, model)

data <- rbind(gamma, gev)

data$model <- factor(data$model, levels = c("Gamma", "GEV"))
data$histogram <- factor(data$histogram, levels = c("Super Naive", "Naive Parametric", "Copula"))

# Generate histograms plot
hist_gamma_gev_FD <- ggplot(data) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(histogram~model, scales = "free_y", ncol = 2)
hist_gamma_gev_FD

ggsave(filename = "hist_gamma_gev_FD.pdf", plot = hist_gamma_gev_FD, path = "../Manuscript", height = 4.2, width = 6)
```

```{r AR(1) D2}
test_copula_D2 <- function(x, pdist, qdist, ..., B) {

    ## Get observed statistic
    n <- length(x) 
    stat  <- ks.test(x, pdist, ...)$statistic

    ## transform the data to standard normal scale and fit arma(1,1)
    zx <- qnorm(rank(x) / (n + 1))

    fit <- tseries::arma(zx, include.intercept = FALSE)
    ar <- fit$coef[1]
    ma <- fit$coef[2]

    ## Perform parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)

    for (i in 1:B) {
        ## 1. The marginal distribution is the fitted dist
        ## 2. There is arma(1,1) dependence on the normal scale
        zx.b <- arima.sim(list(ar = ar, ma = ma), 
                          rand.gen=rnorm,
                          sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)),
                          n = n) # simulate from fit
        u <- pnorm(zx.b)
        # Transform bootstrap sample and fit the parameters
        x.b <- qdist(u, ...)

        # Gather bootstrap sample's test statistic
        stat.b[i] <- ks.test(x.b, pdist, ...)$statistic
    }

    # Return empirical p-value
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

    counter <<- counter + 1
    cat(counter, "")

    return(p.value)
}

# Set variables
n <- 100
ar <- 0.5
ma <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
mean <- 0
sd <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D2(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n), 
                                              pdist = pnorm, 
                                              qdist = qnorm,
                                              mean = mean, 
                                              sd = sd,
                                              B = B))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n), 
                                                pdist = pnorm,
                                                mean = mean, 
                                                sd = sd)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

#save(results, file = "results_ar1_D2.RData")

# Generate histograms plot
hist_ar1_D2 <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_ar1_D2
```

```{r MA(1) D2}
# Set variables
n <- 100
ar <- 0.5
ma <- 0.5
counter <- 0
nrep <- 1000
B <- 100
mean <- 0
sd <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D2(arima.sim(list(ma = ma), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n), 
                                              pdist = pnorm, 
                                              qdist = qnorm,
                                              mean = mean, 
                                              sd = sd,
                                              B = B))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(arima.sim(list(ma = ma), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n), 
                                                pdist = pnorm,
                                                mean = mean, 
                                                sd = sd)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

#save(results, file = "results_ma1_D2.RData")

# Generate histograms plot
hist_ma1_D2 <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_ma1_D2
```

```{r ARMA(1,1) D2}
# Set variables
n <- 100
ar <- 0.5
ma <- 0.5
counter <- 0
nrep <- 1000
B <- 100
mean <- 0
sd <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D2(arima.sim(list(ar = ar, 
                                                           ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n), 
                                              pdist = pnorm, 
                                              qdist = qnorm,
                                              mean = mean, 
                                              sd = sd,
                                              B = B))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(arima.sim(list(ar = ar, 
                                                           ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n), 
                                                pdist = pnorm,
                                                mean = mean, 
                                                sd = sd)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

#save(results, file = "results_arma_D2.RData")

# Generate histograms plot
hist_arma_D2 <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_arma_D2
```

```{r AR(2) D2}
# Set variables
n <- 100
b1 <- 0.5
b2 <- 0.3
counter <- 0
nrep <- 1000
B <- 100
mean <- 0
sd <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D2(arima.sim(list(order = c(2, 0, 0), 
                                                           ar = c(b1, b2)),
                                                      rand.gen=rnorm,
                                                      sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)),
                                                      n = n), 
                                              pdist = pnorm, 
                                              qdist = qnorm,
                                              mean = mean, 
                                              sd = sd,
                                              B = B))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(arima.sim(list(order = c(2, 0, 0), 
                                                           ar = c(b1, b2)),
                                                      rand.gen=rnorm,
                                                      sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)),
                                                      n = n), 
                                                pdist = pnorm,
                                                mean = mean, 
                                                sd = sd)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

#save(results, file = "results_ar2_D2.RData")

# Generate histograms plot
hist_ar2_D2 <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_ar2_D2
```

```{r hist_ar1_ma1_arma_ar2_D2}
load('results_ar1_D2.RData')
model <- rep("AR(1)", 2000)
ar1 <- data_frame(results, model)

load('results_ma1_D2.RData')
model <- rep("MA(1)", 2000)
ma1 <- data_frame(results, model)

load('results_arma_D2.RData')
model <- rep("ARMA(1, 1)", 2000)
arma <- data_frame(results, model)

load('results_ar2_D2.RData')
model <- rep("AR(2)", 2000)
ar2 <- data_frame(results, model)

data <- rbind(ar1, ma1, arma, ar2)

data$model <- factor(data$model, levels = c("AR(1)", "MA(1)", "ARMA(1, 1)", "AR(2)"))
data$histogram <- factor(data$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_ar1_ma1_arma_ar2_D2 <- ggplot(data) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(histogram~model, scales = "free_y", ncol = 4)
hist_ar1_ma1_arma_ar2_D2

ggsave(filename = "hist_ar1_ma1_arma_ar2_D2.pdf", plot = hist_ar1_ma1_arma_ar2_D2, path = "../Manuscript", height = 4.2, width = 6)
```

```{r Gamma D2}
# Set variables
n <- 100
rho <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
shape <- 3
scale <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D2(genData(n, qgamma, shape = shape, scale = scale, copula = normalCopula, rho = rho), 
                                             pdist = pgamma, 
                                             qdist = qgamma,
                                             shape = shape,
                                             scale = scale,
                                             B = B))
counter <- 0

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_standard_ks(genData(n, qgamma, shape = shape, scale = scale, copula = normalCopula, rho = rho), 
                                                pdist = pgamma,
                                                shape = shape,
                                                scale = scale)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

save(results, file = "results_gamma_D2.RData")

# Generate histograms plot
hist_gamma_D2 <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_gamma_D2
```

```{r GEV D2}
# Set variables
n <- 100
rho <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
loc <- 0
shape <- 0.2
scale <- 1

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula_D2(genData(n, qgev, loc = loc, shape = shape, scale = scale, copula = normalCopula, rho = rho),
                                               pdist = pgev,
                                               qdist = qgev,
                                               loc = loc, 
                                               shape = shape,
                                               scale = scale,
                                               B = B))
counter <- 0

# P-vals from standard KS test
pvals_naive <- replicate(nrep, test_standard_ks(genData(n, qgev, loc = loc, shape = shape, scale = scale, copula = normalCopula, rho = rho), 
                                                 pdist = pgev,
                                                 loc = loc, 
                                                 shape = shape,
                                                 scale = scale)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Copula"))

save(results, file = "results_gev_D2.RData")

# Generate histograms plot
hist_gev_D <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_gev_D
```

```{r hist_gamma_gev_D2}
load('results_gamma_D2.RData')
model <- rep("Gamma", 2000)
gamma <- data_frame(results, model)

load('results_gev_D2.RData')
model <- rep("GEV", 2000)
gev <- data_frame(results, model)

data <- rbind(gamma, gev)

data$model <- factor(data$model, levels = c("Gamma", "GEV"))
data$histogram <- factor(data$histogram, levels = c("Naive", "Copula"))

# Generate histograms plot
hist_gamma_gev_D2 <- ggplot(data) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(histogram~model, scales = "free_y")
hist_gamma_gev_D2

ggsave(filename = "hist_gamma_gev_D2.pdf", plot = hist_gamma_gev_D2, path = "../Manuscript", height = 4.2, width = 6)
```

```{r Autoarma}
source("autoarma.R")

# Test non-parametric bootstrap as a solution to fitted parameters
test_autoarma <- function(x, pdist, qdist, B, working) {
  p.value <- ks.arma(x = x, pdist = pdist, qdist = qdist, B = 1000, working = working)
  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

# Naive function that incorrectly uses fitted parameters for gamma distribution
test_super_naive_autoarma <- function(x) {

  m <- mean(x)
  v <- var(x)
  scale <- v/m
  shape <- m*m/v

  p.value <- ks.test(x, "pgamma", shape = shape, scale = scale)$p.value
  list(p.value = p.value)
}

# Generate results
n_tests = 1000
n = 200
phi <- .3
theta <- .2
qdist <- function(u) qgamma(u, shape = 3, scale = 1)
pdist <- function(u) pgamma(u, shape = 3, scale = 1)
working <- "auto"
counter <- 0
pvals_autoarma <- replicate(n_tests, test_autoarma(x = genData(n, phi, theta, qdist), pdist = pdist, qdist = qdist, B = 1000, working = working)$p.value)
counter <- 0
pvals_naive <- replicate(n_tests, test_super_naive_autoarma(genData(n, phi, theta, qdist))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Autoarma", n_tests), rep("Naive", n_tests))
p_values <- c(pvals_autoarma, pvals_naive)
results <- data_frame(p_values, histogram)

results$histogram <- factor(results$histogram, levels = c("Naive", "Autoarma"))

# Generate histograms plot
hist_autoarma <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_autoarma

# Save the histogram figure
ggsave(filename = "hist_autoarma.pdf", plot = hist_autoarma, path = "../Manuscript", height = 2.1, width = 6)
```