### 2023-04-20 manuscript/manuscript.tex
fixed AR(1) confusion in Section 4

\documentclass[12pt, letterpaper, titlepage]{article}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{setspace}

\usepackage[pagewise]{lineno}
%\linenumbers*[1]
% %% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
 \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
 \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
 \renewenvironment{#1}%
 {\linenomath\csname old#1\endcsname}%
 {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
 \patchAmsMathEnvironmentForLineno{#1}%
 \patchAmsMathEnvironmentForLineno{#1*}}%

\AtBeginDocument{%
 \patchBothAmsMathEnvironmentsForLineno{equation}%
 \patchBothAmsMathEnvironmentsForLineno{align}%
 \patchBothAmsMathEnvironmentsForLineno{flalign}%
 \patchBothAmsMathEnvironmentsForLineno{alignat}%
 \patchBothAmsMathEnvironmentsForLineno{gather}%
 \patchBothAmsMathEnvironmentsForLineno{multline}%
}

% control floats
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

\newcommand{\jy}[1]{\textcolor{blue}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{red}{EDS: (#1)}}

\title{On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit}

\author{Anthony Zeimbekakis\\
%   \href{mailto:anthony.zeimbekakis@uconn.edu}
% {\nolinkurl{anthony.zeimbekakis@uconn.edu}}\\
  Elizabeth Schifano\\
  Jun Yan\\[1ex]
  Department of Statistics, University of Connecticut\\
}
\date{}

\begin{document}
\maketitle

\doublespace

\begin{abstract}
The Kolmogorov--Smirnov (KS) test is one of the most popular goodness-of-fit
tests for comparing a sample with a hypothesized continuous distribution.
Nevertheless, it has often been misused. The standard one-sample KS test applies
to independent, continuous data with a hypothesized distribution that is
completely specified. It is not uncommon, however, to see in the literature that
it was applied to dependent, discrete, or rounded data, with hypothesized
distributions containing estimated parameters. For example, it has been
``discovered'' multiple times that the test is too conservative when the 
hypothesized distribution has parameters that need to be estimated.
We demonstrate misuses of the one-sample KS test in
three scenarios through simulation studies:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
For each scenario, we provide remedies for practical applications.

\bigskip
\noindent{\sc Keywords}:
nonparametric bootstrap;
parametric bootstrap;
working dependence.
\end{abstract}

\section{Introduction}
\label{sec:intro}

The Kolmogorov-Smirnov (KS) test is one of the most popular goodness-of-fit
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample of size~$n$ from a continuous
distribution. The null hypothesis $H_0$ is that $X_i$'s follow distribution~$F$.
Let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function. The KS test statistic is
\begin{equation}
  \label{eq:ks_standard}
  D_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The asymptotic distribution of $D_n$ under $H_0$ is independent of the
distribution~$F$. As $n \to \infty$, $D_n$ converges in distribution to
the supremum of standard Brownian bridge \citep{kolmogorov1933sulla}. For large
samples, the tests can be performed with a table \citep{smirnov1948table}.
Critical values for small samples ($n \le 35$) have also been given
\citep{massey1951kolmogorov}. The KS test is available in popular
statistical software packages, such as function \texttt{ks.test()} in R package
\textsf{stats} \citep{R, marsaglia2003evaluating}.

The standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified. In practice, however, it
has often been applied without realizing that one or more of these assumptions
do not hold. For example, \citet{noether1963note} showed the conservativeness of
the KS test when applied to discontinuous distributions. The null distribution
of the KS statistic is no longer distribution free. Computing the
exact and asymptotic distribution of $D_n$ is challenging. Fortunately, the null
distribution of the KS statistic for discontinuous distributions has been
efficiently addressed by \citet{dimitrova2020computing} with a
companion R package \textsf{KSgeneral}. Although a common misuse of KS test, the
issue with discontinuous data is not our focus.

When the hypothesized distribution~$F$ contains unspecified parameters, as is
the case in most goodness-of-fit test settings, the standard KS test is not
applicable. \citet{steinskog2007cautionary} ``discovers'' the change in power
when using fitted parameters and stresses caution in using the KS test in
such ways. In fact, using fitted parameters in place of the true parameters in
KS test has been long known to yield extremely conservative results
\citep[e.g.,][]{lilliefors1967kolmogorov}. This problem can be solved by
parametric bootstrap \citep{efron1985bootstrap, hall1991two}, where
bootstrap samples of the test statistics are constructed from samples
generated from the fitted hypothesized distribution. A nonparametric bootstrap
solution is not trivial because a nonparametric bootstrap sample of the
observed data has ties, which would not happen for continuous distributions.
\citet{babu2004goodness} derived the bias of the standard nonparametric 
bootstrap and
showed how to correct it. They further noted that both parametric and
nonparametric procedures lead to correct asymptotic levels.

The standard KS test does not apply to stationary yet dependent data either. The
distribution of the KS statistic would have a higher variance for positively
dependent data than that derived when the data are independent because of a
smaller effective sample size. For example, for testing normality,
\citet{durilleul1992lack} demonstrates that a naive application of
the KS statistic is too liberal for medium-to-high positive serial dependence,
and that for negative dependence, the behavior is asymmetrical. For remedies,
\citet{weiss1978modification} provides a procedure that is applicable 
specifically for data
modeled by the second-order auto-regressive (AR) process where the AR parameters
are known. \citet{lanzante2021testing} tests various strategies for dealing with
temporal
dependence and concludes that a test based on Monte-Carlo simulations performed
the best. When additionally the hypothesized distribution contains unknown
parameters, the standard KS test becomes even further
inapplicable. Practitioners, however, may have used the test without knowing its
limitations \citep[e.g.,][]{tuncer2019online}.

The contribution of this paper is a demonstration of misuses of the one-sample
KS test in three scenarios and their remedies in practice. The scenarios are
where:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
In each scenario, the misuse is performed and the impacts are shown. Then, a
remedy is detailed and performed alongside the misuse to show its positive
effects. Specifically, unspecified parameters are handled by parametric
bootstrap; serial dependence is handled by introducing a working autoregressive
moving average (ARMA) model that preserves the serial dependence for a wide
range of dependence structures.
In order to set up the demonstrations, simulated data are used
throughout. The remedies are also performed on various families of
distributions. An R implementation of the proposed methods will be available
after the paper is published.

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of
the empirical distribution are serially dependent. A bootstrap procedure
employing a working ARMA model to account for dependence is proposed as a
working solution. Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs. An  
adjusted bootstrap procedure is proposed as a working solution in this case.  
Section~\ref{sec:conclusion} concludes with a discussion.

\section{Unspecified Parameters}
\label{sec:fitted}

The null hypothesis of a goodness-of-fit test is often a composite hypothesis
instead of a single hypothesis. That is, the hypothesized distribution is a
family of distributions with unspecified parameters instead of a specific member
in this family. Let $F_\theta$ be a family of distributions indexed by parameter
vector~$\theta$. The null hypothesis is
\[
  H_0: \text{the random sample $X_1, \ldots, X_n$
    comes from a distribution $F_\theta$ for some $\theta$.}
\]
Since $\theta$ is unknown, one would naturally estimate $\theta$ by
$\hat\theta_n$ from, for example, maximum likelihood or methods of moments. The
KS statistic would then be computed as
\begin{equation}
  \label{eq:ks_fitted}
  D_n = \sqrt{n} \sup_x | F_n(x) - F_{\hat\theta_n}(x) |.
\end{equation}
An overly large $D_n$ still indicates evidence to reject~$H_0$. To get the
p-value of the observed $D_n$, however, note that the null distribution is not
the same as that in the standard
case~\eqref{eq:ks_standard}. If the same null distribution were used, one would
be testing a different~$H_0$ that the random sample comes from the specific
member distribution~$F_{\hat\theta_n}$ instead of the family~$F_\theta$.

The consequence of using the wrong null distribution for the KS statistic $D_n$
in~\eqref{eq:ks_fitted} can be illustrated through a simple simulation study. A
random sample $X_1, \ldots, X_n$ was generated from a normal
distribution with both mean and variance parameters set equal to 8 ($N(8,8)$), 
and sample size $n = 200$. The maximum likelihood estimates of the mean and 
variance were used as $\hat\theta_n$. The p-value of the test
statistic~$D_n$ in Equation~\eqref{eq:ks_fitted} was obtained by naively calling 
the \texttt{ks.test()} function in~R with the fitted normal distribution as the
hypothesized distribution. That is, the hypothesized distribution was
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample
variance (with $n$ in the denominator). This experiment was repeated 1000 times
and the probability-probability plot (PP-plot) of the 1000 p-values are 
displayed in the top left plot of Figure~\ref{fig:pp_f}. If the test were
valid, the p-values would be uniformly distributed in the unit
interval and the points in the PP plot would fall along the diagonal line. 
Clearly, the distribution of the 1000 p-values are very different from what
one would expect to see from 1000 draws from the standard uniform
distribution.

We similarly considered random samples $X_1, \ldots, X_n$ 
generated from a gamma distribution with shape and scale parameters 8 and 1, 
respectively ($\Gamma(8,1)$), and sample size $n=200$. These parameter values 
were selected so that mean and variance of the normal and gamma distributions 
in the simulations are the same. Likewise, the p-value of the test 
statistic~$D_n$ in~\eqref{eq:ks_fitted} was obtained by naively calling the
\texttt{ks.test()} function in~R with the fitted gamma distribution as the
hypothesized distribution using the shape and scale maximum likelihood 
estimates. The PP plot of the p-values from 1000 replications of this 
experiment are displayed in the bottom left plot of Figure~\ref{fig:pp_f}. The
p-values are again non-uniformly distributed due to the use of the 
estimated parameters when specifying the null distribution.

To fix the problem, parametric bootstrap can be used to approximate the null
distribution of the test statistic~$D_n$:
\begin{enumerate}
\item
  Draw a random sample $X_1^*,...,X_n^*$ from the fitted distribution
  $F_{\hat\theta_n}$
\item
  Fit $F_\theta$ to $X_1^*,...,X_n^*$ and obtain estimate 
	$\hat\theta_n^*$ of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of
  $X_1^*, \ldots,  X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sqrt{n} \sup_x | F_n^* (x)- F_{\hat\theta_n^*}(x) |.
  \]
\item
  Repeat the previous steps a large number~$B$ times and use the empirical
  distribution of $D_n^*$ to approximate the null distribution of the observed
  statistic.
\end{enumerate}
The p-value of $D_n$ is approximated by the proportion of the $D_n^*$ 
statistics that are
greater than or equal to~$D_n$.

In the same simulation study, p-values were obtained for the 1000 replicates
from the parametric bootstrap procedures. The top and bottom right plots of 
Figure~\ref{fig:pp_f} display the PP-plots of the 1000
p-values from the normal and gamma data generation scenarios,
respectively. This time, all points fall along the diagonal lines implying that
the p-values are coming from standard uniform distributions.

Nonparametric bootstrap can also be used to approximate the null distribution
of the test statistic in~\eqref{eq:ks_fitted} except that a bias correction is 
needed
\citep{babu2004goodness}. Details of the nonparametric bootstrap procedure are
in the Appendix. The results from nonparametric bootstrap are similar to those
from the parametric bootstrap and, hence, are omitted.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.5\textwidth]{pp_f.pdf}
  \caption{PP-plots for the p-values obtained from KS tests with unspecified
    parameters with sample size $n = 200$ based on 1000 replicates. The data
    generating models were $N(8, 8)$ and $\Gamma(8, 1)$.
  }
  \label{fig:pp_f}
\end{figure}

\section{Serially Dependent Data}
\label{sec:dependence}

Here we consider the situation where the observed data $X_1, \ldots, X_n$ are
not independent but serially dependent, and the hypothesized distribution $F$
contains no unknown parameters. That is,
\[
H_0: \text{$X_i$'s have marginal distribution $F$.}
\]
The KS statistic $D_n$ in Equation~\eqref{eq:ks_standard}
is still a good measure of deviation from the null hypothesis. Nonetheless, the
null distribution of $F_n$ changes when there is serial dependence.
When the serial dependence is
positive, the effective sample size is less than~$n$. The test statistic
would be stochastically greater than that obtained from independent data, 
resulting in a test that is too liberal (i.e., the null hypothesis is rejected
more often than it should). The reasoning is similar when the serial dependence 
is negative, in which case, the test is conservative.

The invalidity of the standard statistic for serially dependent data can be
illustrated by a simple simulation study. Consider a first-order autoregressive
or AR(1) model with marginal normal (or gamma) distribution. For each AR
coefficient $\psi \in \{-0.3, -0.1, 0.1, 0.3\}$, we generated 1000
series of length 200. For each series, we applied the KS statistic
in~\eqref{eq:ks_standard} to test $H_0$ that the $X_i$'s follow $N(8, 8)$ (or 
$\Gamma(8,1)$) marginally. The PP-plots of the p-values from $1000$ replicates 
for all the $\psi$ values are displayed in Figure~\ref{fig:pp_s}. As 
$\phi$ diverges from 0, the p-values look less likely to be generated from a 
standard uniform distributions. Positive $\psi$ values led to p-values with a 
distribution function higher than the standard uniform distribution function and, 
hence, conservative tests. Negative $\psi$ values led to p-values with a 
distribution function lower than the standard uniform distribution and, hence, 
liberal tests. \eds{the previous two sentences are not consistent with what is 
written at the beginning of this section. If I am interpreting the plots 
correctly, then positive $\psi$ leads to p-values that are too low which would
correspond to liberal tests (consistent with what was written at the beginning 
of this section).}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{pp_s.pdf}
  \caption{PP-plots for the p-values obtained from KS test with serial
    dependence ignored and sample size $n = 200$ based on 1000
    replicates. The data have marginal distributions $N(8, 8)$ and
    $\Gamma(8, 1)$, with an AR(1) dependence structure on the standard normal
    scale. The AR coefficients were $\psi \in \{-0.3, -0.1,  0.1,  0.3\}$.
  }
  \label{fig:pp_s}
\end{figure}

A complete remedy for KS tests with serially dependent data is
challenging. The null distribution depends on the structure of the serial
dependence, which can be arbitrary in practice. A parametric bootstrap procedure
would need to specify a model for the dependence, whereas the primary interest
is to test the marginal distribution of the stationary series. A block
nonparametric bootstrap for stationary data \citep{kunsch1989jackknife} is
tempting, but the counterpart of a bias correction as in
\citet{babu2004goodness} is not available yet. Is there any approach that does
not require full specification of the dependence of the stationary process,
but still gives reasonably satisfactory correction to the size of the KS test in
certain applications?

We propose a semiparametric bootstrap procedure that assumes a working serial
dependence structure which does not need to be correctly specified. The working
serial dependence is introduced through a working ARMA process with the hope to
approximate the true serial dependence as allowed by the working
model. Essentially, this working model covers the most commonly seen dependence
structure characterized by a normal copula \citep{hofert2018elements}, which
separates the dependence structure
of a multivariate distribution from its marginal distributions. 
The parameters of the working normal copula are
estimated from fitting an ARMA model to the
observed data transformed to the standard normal scale.

In particular, let
$Z_i = \Phi^{-1}\{ F(X_i)\}$, $i = 1, \ldots, n$, where $\Phi$ is the distribution
function of the standard normal. Then we fit an ARMA$(p, q)$ model to
$Z_1, \ldots, Z_n$ with AR order $p$ and MA order $q$ selected by the Akaike
Information Criterion (AIC). This can be done with, for example, function
\texttt{auto.arima()} from  R package \texttt{forecast}
\citep{hyndman2008automatic}. Since it is known that the mean of $Z_i$'s is
zero, it is necessary to set the intercept of the ARMA model as zero in the
fitting process. This restriction turned out to be critical in our
investigation; having the intercept estimated does not lead to desired p-value
distributions in the following bootstrap process.
The selected ARMA$(p, q)$ model with fitted
parameters will be used as the working model to generate bootstrap samples with
serial dependence mimicking that in the observed data. The unconditional
variance $\sigma^2$ of the ARMA model with unit innovation variance can be
obtained with function
\texttt{tacvfARMA()} from R package \texttt{ltsa} \citep{mcleod2007algorithms}.

The semiparametric bootstrap procedure is as follows.
\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the fitted ARMA$(p, q)$ process with
  innovation variance $1 / \sigma^2$ such that the $Z_i$'s are
	marginally standard normal variables.
\item
  Form a bootstrap sample $X_i^* = F^{-1} [\Phi(Z_i^*)]$,  $i = 1, \ldots, n$,
  where $F^{-1}$ is the quantile function of $F$.
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x \lvert F_n^* (x)- F(x) \rvert.
  \]
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of the $B$ test statistics to approximate
  the null distribution of the observed statistic.
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$.

This process is semiparametric because the marginal distribution is
parametrically specified while the dependence structure of the stationary series
is only approximated by that of an ARMA process with normal distributions.
The closer the approximation is to the truth, the better
performance of the size of the KS test. The normal copula of the working ARMA
process provides a wide class of dependence structures.
If the true dependence indeed has an ARMA structure, this method is exact. When
the true dependence is not covered by the ARMA model, the working model may 
still give a reasonable approximation that can
be useful for practical purposes.

The semiparametric bootstrap procedure was evaluated in simulation. We 
considered three true dependence structures on the standard normal scale 
(ARMA(1, 1), AR(2), and MA(1)) and two marginal distributions 
($N(8,8)$ and $\Gamma(8,1)$).  For each dependence structure and marginal distribution combination, we generated 1000 series of length 200.  For each series, we applied the semiparametric bootstrap procedure with serial
dependence accounted for through a working ARMA model to test $H_0$ that the 
$X_i$'s marginally follow their true marginal distribution ($N(8, 8)$ or 
$\Gamma(8,1)$). Figure~\ref{fig:pp_ss} displays the PP-plots of the 1000 
p-values for each dependence structure and marginal distribution combination.
All plots within the figure show the points falling along the diagonal lines, 
implying that the p-values are coming from standard uniform distributions and 
that the semiparametric bootstrap procedure is an effective remedy in all cases
considered.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.75\textwidth]{pp_ss.pdf}
  \caption{PP-plots for the p-values obtained from KS test with serial
    dependence accounted for through a working ARMA model and $n = 200$ based on
    1000 replicates. The data have marginal distributions $N(8, 8)$ and
    $\Gamma(8, 1)$. The dependence structure on the standard normal scale was
    characterized by ARMA(1, 1) model, AR(2) model, and MA(1) model.
  }
  \label{fig:pp_ss}
\end{figure}

\section{Unspecified Parameters and Serially Dependent Data}
\label{sec:fittedwithdependence}

When the observed sample $X_1, \ldots, X_n$ is no longer a random sample but
a serially dependent stationary series, we may still be interested in testing
whether marginally each $X_i$ follows hypothesized distribution with unknown
parameters. That is, the null hypothesis is
\[
H_0: \text{$X_i$'s have marginal distribution $F_\theta$ for some $\theta$}.
\]
The testing statistic $D_n$ in Equation~\eqref{eq:ks_fitted} still measures the
deviation from $H_0$, but its null distribution depends on the serial
dependence. The bootstrap procedure in the last section can be adapted to handle
this situation.

Specifically, let $\hat\theta_n$ be the marginally fitted parameters, and
$Z_i = \Phi^{-1} \{F_{\hat\theta_n}(X_i)\}$, $i = 1, \ldots, n$ be a
transformation of the observed data onto the standard normal scale using the
fitted parameters $\hat\theta_n$. We then fit an ARMA model with orders
selected by the AIC using function \texttt{auto.arima()} from R package
\texttt{forecast}. This ARMA model will be used as the working model to
introduce serial dependence in the bootstrap samples. Again, we obtain the
unconditional variance $\sigma^2$ of the ARMA process with unit innovation
variance. Our semiparametric bootstrap procedure is as follows.

\begin{enumerate}
\item
  Generate $Z_1^*, \ldots, Z_n^*$ from the working ARMA$(p, q)$ model with
  innovation variance $1 / \sigma^2$ such that the $Z_i$'s are marginally
  standard normal variables.
\item
  Form a bootstrap sample
  $X_i^* = F_{\hat\theta_n}^{-1} [\Phi(Z_i^*)]$, $i = 1, \ldots, n$, where
  $F_{\hat\theta_n}^{-1}$ is the quantile function of $F_{\hat\theta_n}$.
\item
  Fit $F_\theta$ to $X_1^*, \ldots, X_n^*$ to obtain estimate $\hat\theta_n^*$
  of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of the bootstrap sample
  $X_1^*, \ldots, X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x \lvert F_n^* (x) - F_{\hat\theta_n}^* (x) \rvert.
  \]
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of the $B$ test statistics to approximate
  the null distribution of the observed statistic.
\end{enumerate}
The p-value of $D_n$ is approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Compared to the algorithm in
the last section, here we use $F_{\hat\theta_n}$ and $F_{\hat\theta_n}^*$ in
place of the known~$F$.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=.75\textwidth]{pp_s_f.pdf}
  \includegraphics[width=.75\textwidth]{pp_ssf.pdf}
  \caption{PP-plots for the p-values obtained from KS test without (top) and 
	  with (bottom)
    serial dependence accounted for and unspecified parameters and sample size
    $n = 200$ based on
    1000 replicates. The data have marginal distributions $N(8, 8)$ and
    $\Gamma(8, 1)$. The dependence structure on the standard normal scale was
    characterized by ARMA(1, 1) model, AR(2) model, and MA(1) model.
  }
  \label{fig:pp_ssf}
\end{figure}

Figure~\ref{fig:pp_ssf} shows the results of the
procedure performed on data generated with dependence structures of ARMA(1, 1),
AR(2), and MA(1). Naively fitting parameters and not adjusting for dependence  clearly deviates from a uniform distribution of p-values. 
Parametric bootstrap provides some correction but does not account for 
dependence, 
\eds{just want to verify: you are using parametric bootstrap from Section 2 for
the top portion of Figure~\ref{fig:pp_ssf}? (Not the totally naive procedure 
that ignores both issues)} 
so therefore the results of the copula remedy are more accurate and favorable.
%In the case of MA(1) and ARMA(1, 1), the true dependence appears close enough
%to our assumption of AR(1) that the results are reasonable. AR(2) however
%appears to be just far enough from our assumption, showing a limitation of the
%procedure.

Similar to Section~\ref{sec:dependence}, the copula approach is not a complete
solution. Regardless of the true dependence in the data we assume an ARMA$(p,q)$
dependence structure. As long as this assumption is a close approximation, the
correction should still provide a reasonable approximation that can be useful 
for practical purposes. 
%Regardless of the true dependence in the data we assume an AR(1)
%dependence structure by taking the lag-1 sample auto-spearman rho. However, we
%can show that as long as the AR(1) assumption is a close approximation, the
%correction still provides a reasonable approximation that can be useful for
%practical purposes.

\begin{table}[ht]
  \label{tab:power}
  \caption{Power of rejecting the null hypothesis of normal and gamma
    distribution at significance level 0.05 when the true distribution was
    $\Gamma(8, 1)$ and $N(8, 8)$, respectively. }
\centering
\begin{tabular}{llrrrr}
  \toprule
  &  & \multicolumn{2}{c}{$H_0:$ normal} &
  \multicolumn{2}{c}{$H_0$: gamma}\\
  \cmidrule(lr){3-4}\cmidrule(lr){5-6}
  Dependence & Method & $n = 100$ & $n = 200$ & $n = 100$ & $n = 200$\\
  \midrule
ARMA & naive & 0.40 & 0.66 & 0.52 & 0.78 \\ 
   & copula & 0.33 & 0.56 & 0.44 & 0.72 \\ 
  AR & naive & 0.41 & 0.66 & 0.46 & 0.70 \\ 
   & copula & 0.26 & 0.45 & 0.31 & 0.54 \\ 
  MA & naive & 0.37 & 0.73 & 0.51 & 0.82 \\ 
   & copula & 0.34 & 0.69 & 0.49 & 0.80 \\ 
   \bottomrule
\end{tabular}
\end{table}

Finally, we report a small study of the power of the one-sample KS test with
unspecified parameters and serially dependent data. Using the three dependence
models characterized by ARMA(1, 1), AR(2), and MA(1), we generated serially
dependent data with marginal distributions $N(8, 8)$ left truncated by zero and
$\Gamma(8, 1)$. Then, for data from $\Gamma(8, 1)$, we tested $H_0$ that the
data has marginally a normal distribution; for data from $N(8, 8)$ left
truncated by zero, we tested $H_0$ that the data has marginally a gamma
distribution;  In both cases, no parameters were specified. Two sample sizes
were considered, $n \in \{100, 200\}$. Table~\ref{tab:power} summarizes the
power of rejecting the null hypothesis at significance level~0.05. The tests
based on a working copula have substantial power in rejecting the null
hypothesis and the power increases as sample size increases. The power from the
naive tests where the serial dependence is ignored, although the unspecified
parameters are accounted for \eds{i.e., procedure from Section 2?  It may be 
less confusing if we had a name for the procedure discussed in each section 
rather than using ``naive'' to describe both levels of ignorance?}, appear to 
have higher power than the proposed tests, but this is not reliable since they 
do not hold their sizes as demonstrated in the Figures.

\section{Conclusion}
\label{sec:conclusion}

The KS test has base assumptions that the hypothesized distribution is
completely specified and the data is independent. When these assumptions are
violated, the test is no longer accurate and remedies must be performed. In the
case of fitted parameters, parametric and non-parametric bootstrap can restore
the size of the test. A bias correction is required if the non-parametric form
is used \citep{babu2004goodness}. In the case of serially dependent data, 
a procedure using bootstrap with copulas to model dependency shows positive
results. When both assumptions are violated, i.e., where the data has serial 
dependence and parameters must be fitted, an adjusted copula procedure also 
shows positive results. The tests appear effective for a variety of families of 
distributions. The copula remedy is not a complete solution and has limitations. 
Regardless of the true dependence, we assume an ARMA($p,q$) dependence structure. 
%Therefore, if the AR(1) dependence structure is a 
%close approximation of the truth, the approach can work.
%However, in cases such as AR(2), if 
%the approximation is too far from the true
%dependence structure the approach does not completely remedy the issue.
Therefore, if the assumed dependence structure is a close approximation of the
truth, the approach can work. However, if the approximation is too far from the 
true dependence structure the approach does not completely remedy the issue. 
As well, tests were only performed with the normal copula. It is possible that 
other copulas could provide stronger results based on the true dependence of the 
data.

\appendix

\section{Nonparametric Bootstrap for KS Test with Fitted Parameters}

Using the same notations in the text, the procedure is summarized as follows
\citep{babu2004goodness}.
\begin{enumerate}
\item
  Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
  with replacement
\item
  Fit $F_\theta$ to $X_1^*,...,X_n^*$ and obtain estimate 
	$\hat\theta_n^*$ of $\theta$.
\item
  Obtain the empirical distribution function $F_n^*$ of
  $X_1^*, \ldots,  X_n^*$. 
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sup_x | \sqrt{n}\left(F_n^* (x)- F_{\hat\theta_n^*}(x)\right) - B_n(x) |.
  \]
  where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known
  bias term \citep{babu2004goodness}.
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of $D_n^*$ to approximate the null distribution of the observed
  statistic. 
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of the $D_n^*$
statistics that are greater than or equal to~$D_n$. Note that the step~1 and
step~4 are different from the parametric bootstrap version.

\bibliographystyle{asa}
\bibliography{citations}

\end{document}
%%% LocalWords: 
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% ispell-personal-dictionary: ".aspell.en.pws"
%%% fill-column: 80
%%% eval: (auto-fill-mode 1)
%%% End: