### 2024-04-16 Code/UG_KStest_Assignment.Rmd
supplemental UG assignment added

---
title: "Misuse of Kolmogorov-Smirnov One-Sample Goodness of Fit Test"
author: "STAT3215"
date: "Fall 2023"
output:
  pdf_document:
    extra_dependencies:
    - amsmath
    - amssymb
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The Kolmogorov-Smirnov (KS) test is one of the most popular goodness-of-fit
tests for comparing a sample with a hypothesized parametric distribution. 
We used this test earlier this semester (Note 5) to assess whether the 
(externally) studentized residuals from multiple linear regression can be viewed 
as realizations of a standard normal distribution ($N(0,1)$) for large samples 
($n-p \gg 30$, where $n$ is the sample size and $p$ is the number of regression
parameters in the mean function).

Despite its popularity, the test is frequently misused in the literature and
practice. While originally intended for independent, continuous data with
precisely specified hypothesized distributions, it is erroneously applied to
scenarios with dependent, discrete, or rounded data, with hypothesized
distributions requiring estimated parameters. For example, it has been
``discovered'' multiple times that the test is too conservative when the 
hypothesized distribution has parameters that need to be estimated.  We are 
going to explore this scenario in more detail below.

## KS Test Details

Let $X_1, ..., X_n$ be a random sample of size $n$ from a continuous
distribution. The null hypothesis $H_0$ is that $X_i$'s follow distribution $F$.
Let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function. The KS test statistic is
$$D_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.$$
The asymptotic distribution of $D_n$ under $H_0$ is independent of the
distribution $F$.  Critical values have been tabulated for large and 
small samples alike with popular statistical software packages readily able to 
provide associated p-values. In particular, the KS test is available as function 
\texttt{ks.test()} in R package \textsf{stats}.

This standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified (meaning the parameters
of the hypothesized distribution are known and not estimated). When the 
hypothesized distribution $F$ contains unspecified parameters, as is
the case in most goodness-of-fit test settings, the standard KS test is not
applicable.  We will illustrate this below with a small simulation study.

## Illustration

### Simulation when KS test conditions are met

We first consider a simulation when the KS test conditions are met, namely, that 
the hypothesized distribution of independent data is continuous and completely 
specified. In this illustration, the hypothesized distribution $F$ is a normal
distribution with both mean and variance parameters set equal to 8 ($N(8,8)$).

We will evaluate the performance of the standard KS test when the null hypothesis 
is true by repeating the following steps $B=10000$ times:

1. Generate a sample of $n=200$ independent observations from a $N(8,8)$ distribution.

2. Perform the standard KS test on the sample generated in Step 1 by calling the
\texttt{ks.test()} function with known hypothesized distribution $N(8,8)$.

3. Save and store the p-value reported from Step 2.

We run these steps in the code below and then plot a histogram of the resulting $B=10000$ p-values.

```{r, fig.height = 3, fig.width = 3.5, fig.align = "center", fig.cap="\\label{fig:hist}Histogram of p-values when KS test conditions are met."}
do1rep <- function(n) {
    x <- rnorm(n, mean = 8, sd = sqrt(8))
    ks <- ks.test(x, "pnorm", mean = 8, sd = sqrt(8))
    ks$p.value
}

n <- 200   # sample size
B <- 10000 # number of simulation replicates

set.seed(0)  # for reproducibility
pvals <- replicate(B, do1rep(n))

hist(pvals, main="")
```

Recall that for each replication, we generated the data so that the null 
hypothesis of the KS test is true. Because the test is being performed under the 
appropriate conditions, the test should be valid in the sense that the resulting 
p-values should have a continuous uniform distribution on the unit interval 
($U(0,1)$). Indeed, the histogram in Figure \ref{fig:hist} reflects this. This 
also means that if we set the significance level at $\alpha=0.05$ so that we
reject $H_0$ when p-value$<\alpha$, this false rejection should happen in 
approximately 5\% of the $B=10000$ simulations. We compute the proportion of 
false rejections from this simulation with the code below.  
```{r}
mean(pvals<0.05)
```

### Assignment: Simulation when KS Conditions are NOT met

We will now consider a simulation when the KS test conditions are NOT met, 
namely, that the hypothesized distribution of independent data is continuous, 
but the parameters of the hypothesized distribution are unknown to the 
researcher, and are replaced with estimated values.  Your assignment is to 
evaluate the performance of the standard KS test under this scenario.

Specifically, you will modify the simulation code from the previous section to 
perform the following four steps, repeated $B=10000$ times:

1. Generate a sample of $n=200$ independent observations from a $N(8,8)$ 
distribution.

2. Estimate the sample mean, $\bar{x}$ and the sample variance, $s^2$, of the sample 
generated in Step 1 using the \texttt{mean()} and \texttt{var()} functions, 
respectively.

3. Perform the standard KS test on the sample generated in Step 1 by calling the
\texttt{ks.test()} function with hypothesized distribution $N(\bar{x},s^2)$.

4. Save and store the p-value reported from Step 3.

After repeating Steps 1-4 $B=10000$ times, plot a histogram of the p-values. If 
the test remains valid when using estimated parameters in the hypothesized 
distribution, we would expect the p-values to uniformly distribute on the unit 
interval as they did in the first simulation (Figure \ref{fig:hist}) and that 
the probability of false rejection of the null hypothesis is the significance 
level $\alpha$.

Do the p-values appear uniformly distributed in the unit interval here, or is 
the distribution skewed to the left or right?

Report the proportion of false rejections in this simulation using significance 
level $\alpha=0.05.$