### 2024-04-29 manuscript/TAS_cover_revision.tex
add cover, minor typo fixes to ms and reply

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
%\topmargin=-0.0in    % Make letterhead start about 1 inch from top of page
%\textheight=8.5in  % text height can be bigger for a longer letter
%\oddsidemargin=0pt % leftmargin is 1 inch
%\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\usepackage[margin=1in]{geometry}

\begin{document}

\signature{Elizabeth D. Schifano \\elizabeth.schifano@uconn.edu \\Corresponding Author}           % name for signature
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{

\textit{The American Statistician} Manuscript Revision}

\begin{flushleft}
{\large\bf Elizabeth D. Schifano, Ph.D.}
\end{flushleft}
\medskip\hrule height 1pt
\begin{flushright}
\hfill University of Connecticut \\
\hfill Department of Statistics \\ 
\hfill 215 Glenbrook Road U-4120 \\
\hfill Storrs, CT 06269
\end{flushright}
\vfill % forces letterhead to top of page

%\vspace{-2in}
\opening{Dear members of the \textit{The American Statistician} editorial team:}
\vspace{.25in}

\noindent Thank you for the opportunity to revise and resubmit the manuscript 
entitled ``On Misuses of the Kolmogorov–Smirnov Test 
for One-Sample Goodness-of-Fit'' (TAS-23-210) for possible publication in 
\emph{The American Statistician}.\\

We have now included a new section in the revised manuscript, \textit{Section 5 
Classroom Implementation}, that discusses how elements of this manuscript have 
been incorporated into both graduate- and undergraduate-level teaching at the 
authors' home university.

We have also included Supplemental Materials that include an \texttt{R} 
implementation of the proposed methods, as well as teaching materials at the 
undergraduate and graduate level. \\

We hope that you continue to find this manuscript to be a valuable contribution 
to the Teacher's Corner in the \emph{The American Statistician}. Please inform 
me if I can provide any additional information or assistance.  
\\

Thank you very much for your continued consideration. I look forward to hearing 
from you.\\

\closing{Best regards,}

~~\\
%\encl{Online Updating of Statistical Inference in the Big Data Setting\\~~~ Supplemental Material}  				% Enclosures

\end{letter}

\end{document}